---
title: "HW4"
author: "Wilson Lui (WL2522)"
output:
  html_document: default
  html_notebook: default
---


#1)


Write a DRAFT introduction. (It will likely change by the time you submit the final project.) From the final project assignment: 

"In this section, explain why you chose this topic, and the questions you are interested in studying. Include a brief description of how you found the data, and clear instructions on where the reader can find the data."


##Introduction


> This project is an exploratory data analysis of a dataset originating from the computer game "Starcraft II" (henceforth referred to as "SC2" for short). SC2 belongs to the real-time strategy (RTS) genre, which typically involve players managing military bases and directly controlling armies in order to combat opposing players' armies and destroy their bases. All player actions happen in real-time. Thus, a successful player must be able to construct buildings, gather resources, and raise an army while simultaneously controlling available combat units to attack and defend. This dichotomy is referred to macromanagement and micromanagement, or "macro" and "micro" for short. A more detailed description of the game written by the researchers who compiled this dataset can be found [here](http://skillcraft.ca/about-the-project.html).


> The dataset used for this analysis originates from the research paper [Video Game Telemetry as a Critical Tool in the Study of Complex Skill Learning](http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0075129#pone-0075129-g001) (Thompson et al, 2013) and can be found [here](http://summit.sfu.ca/item/13328). From August 12th, 2011 to September 19th, 2011, the researchers solicited members of various online SC2 community hubs for their game replay files, from which I found out about the study. The purpose of this study is to explore SC2 as a tool for measuring how people learn and improve skills. These replay files are an exact record of everything that happens during the course of a game, including various metrics related to the game state as well as the locations of every game unit at any given point in time (the video previously referenced was made using one such a replay file). These games were played on the 1 vs. 1 online ladder system, which places players in one of seven ladder ranks according to skill level. One's ladder rank is based on his or her win-loss record and the ranks of both players at the time of each game. When a player wishes to play, the system attempts to find an opponent with approximately the same skill level. 


> What makes SC2 an interesting game to study is that the sheer complexity of the game in terms of all the things that could happen at any given moment compounded across an entire game means that every game is unique. The popularity of SC2 has led to a sustained professional scene, where players are sponsored by teams and compete in tournaments for cash prizes. These tournaments might be played online or offline, with players being flown into a studio or arena with a live audience. These events are accompanied by commentators who serve a similar role to those in traditional sports. In the past, the firt place prizes at the most prestigious tournaments have reached $100,000. This competitive scene has provided the game's passionate fanbase with a clear motivation for becoming better at the game while these professional players and their serve as role models for how to become better (A beginner's guide to competitive SC2 can be found [here](https://www.youtube.com/watch?v=JSPRgL4D1no)). As a passionate member of this community during the time that this dataset was compiled, I made a serious (but short-lived) effort at improving my SC2 skill level, from which I accumulated various advice and generally accepted wisdom on how one can get better at the game, as well as a deep appreciation for the talents of professional SC2 players. With this dataset, I would like to see if these community best practices are supported by the differentiaton of skill levels based on the ladder rank of the study participants. In their paper, the researchers discussed the next study that they were embarking on, which involved accepting large numbers of replay files from participants in order to assess changes in their skill over time. Though each data point in this dataset represents a single moment in time for one player, I want to explore how in-game metrics change across age groups and experience levels both within and across ranks. I plan on using variables in the dataset related to time to see if there are any general trends over time that can be discerned.


> The researchers approached this study from a cognitive science perspective, introducing a new unit of measurement called the Perception-Action Cycle (PAC), which attempts to quantify skill in terms of the amount of time between perceiving an event and acting upon it. They hypothesized that the variables that are most correlated with skill change as a player improves and moves up the ranks and showed that machine learning methods can be used to predict a player's rank based on such variables. I wish to avoid using variables related to PAC since I don't believe that I have the necessary cognitive science background to understand their meaning and significance. Also, I want to avoid retreading the path the researchers took, since their paper focused heavily on PAC and its implications on SC2 skill level.


#2) 


Write a DRAFT description of the Team section.

"List team members and a description of how each [WILL] contribute to the project."


> Since I am working alone for this project, I am planning on exploring and analyzing this dataset in the following order:


1. Check the self-reported variables for nonsensical/missing/outlier values and decide on a way to deal with them
2. Propose an approximate order of in-game metric importance (in terms what the SC2 community deems essential for increasing skill level as opposed to what's "nice to have", similar to Maslow's heirarchy of needs)
3. Explore these variables individually in the order proposed
4. Visualize these variables together with variables related to time (player age, hours played per week, total hours played) to get an imprecise idea of how these metrics changes over time
5. Create multivariate plots to find possible relationships between different metrics


#3) 


Choose four variables that you consider to be important and begin analyzing them, following the advice in Chapter 8, sections 1 & 2, of Graphical Data Analysis with R, to begin by summarizing and plotting individual variables. Histograms and bar charts work well for this purpose, but you can use whatever techniques you find appropriate. (However, do not use graphs involving circles to communicate information. That includes pie charts, donut charts, polar coordinates, scatterplots using the size of the dots to represent quantities, and similar. We will discuss the reason for this rule in class.)


> My analysis begins with loading the csv file containing the dataset and converting integer variables that were imported as factors. I then create a new variable called *rank* that adds the proper name for each player's league based on the *LeagueIndex* variable.


```{r}

library(tidyverse)

sc <- read.csv('StarCraft2ReplayAnalysis-1.csv')

#Convert factors back to integers
#Automatically replace "NULL" string values with "NA"

sc$TotalHours <- as.integer(as.character(sc$TotalHours))
sc$HoursPerWeek <- as.integer(as.character(sc$HoursPerWeek))
sc$Age <- as.integer(as.character(sc$Age))

#Add a new variable 'rank' so that the values in the included "LeagueIndex" variable are associated with the names of each league

rank <- c('Bronze', 'Silver', 'Gold', 'Platinum', 'Diamond', 'Master', 'GrandMaster', 'Professional')
rank <- factor(rank, levels = c('Bronze', 'Silver', 'Gold', 'Platinum', 'Diamond', 'Master', 'GrandMaster', 'Professional'))
sc <- mutate(sc, League = rank[LeagueIndex])

```


> Since my analysis will partly be focused on the change in player metrics as a function of time, it would be appropriate to isolate the two time variables, *HoursPerWeek* and *"*TotalHours*"*. Like other variables that are outside the scope of any single game, they contain self-reported values based on a survey that participants filled out when submitting their replay files. Thus, they are prone to reporting inaccuracy due to the fact that participants are attempting to generalize several months of playing experience and habits into two numbers. 


```{r}

summary(sc$HoursPerWeek)
summary(sc$TotalHours)

```


> What immediately stands out in the summary statistics is the maximum values for both variables. The maximum value for *HoursPerWeek* corresponds to 24 hours of playing time every week, while the maximum reported value of 1,000,000 for *TotalHours* far exceeds the number of hours that had elapsed since SC2 was released to the time that the data collection period concluded. When I discussed this matter with Professor Robbins, one of her suggestions was to "fence" the values such that outliers are replaced with a sensible ceiling value. From interviews of professional players as well as hearsay from those who are in touch with them, I chose 98 hours played per week as a ceiling value. This corresponds to 14 hours played per day, which is a habit that could most likely only be sustained by professional players signed to a major team. Such teams often operate a "team house" where all the players live and train together. It is not uncommon for the sponsors to hire maids to cook and clean so that the players can focus all their attention on playing SC2. In such cases, it is not unusual for players to devote 10-14 hours per day to practicing for upcoming important matches. As for the single data point with a value of 0 hours per week, I interpreted that as the respondent playing the game very infrequently. Because there was only one such data point and the fact that the next closest value is 2, I decided not to remove it. This variable appears roughly unimodal after adjusting the bandwidth of the density plot, with a mode of approximately 7 and a mean of 15.87. 


```{r}

#Restrict "HoursPerWeek" to a maximum value of 98

sc$HoursPerWeek <- ifelse(sc$HoursPerWeek > 98, 98, sc$HoursPerWeek)

summary(sc$HoursPerWeek)

ggplot(sc, aes(sc$HoursPerWeek)) + geom_density(bw = 2.5) + labs(x = 'Reported Hours Played Per Week') + scale_x_continuous(breaks = c(seq(0, 100, by = 5)))

```


> Similarly, I restricted the maximum value for *TotalHours* to 8,106. I arrived at this theoretical maximum by finding the number of days from the earliest date that a member of the general public could begin playing the game consistently, the start of the pre-release beta test period on February 17th, 2010, to the end of the data collection period, September 19th, 2011. Assuming an extremely dedicated fan played the theoretical maximum of 14 hours per day every day during that period, the played will have played the game for 8,106 hours.


```{r}

sc$TotalHours <- ifelse(sc$TotalHours > 8106, 8106, sc$TotalHours)

ggplot(sc, aes(sc$TotalHours)) + geom_histogram() + labs(x = 'Reported Total Number of Hours Played')

```


> In the above histogram, several outlier values well below 8,106 hours can be seen. 